----
title: 'DEseq_normalization of H3K4me3 CUT&Tag data done in CAST-KI testicular cells
  and checking reproducibility between replicates using Pearson Correlation Coefficient '
author: "Aditya Mahadevan"
date: "1/27/2021"
output:
  html_document: default
  pdf_document: default
----
This is the Rmarkdown document of the analysis carried out to check the reproducibility between H3K4me3 CUT&Tag replicates and also between H3K4me3 Cut&Tag merged and H3K4me3 ChIP merged samples. This data has been generated in CAST-KI testicular cells. 
#Loading all the required packages for the analysis
```{r setup, include=FALSE}
library(tidyverse)
library(edgeR)
library(ggpubr)
```
#First lets use the peakome file generated for CUT&Tag replicates on CUT&Tag merged MACS e-1 peak dataset
#Loading the file and then converting into a dataframe to look at the plot before normalization
```{r cars}
df <- as.matrix(read.table("CUTnTag_rep1_rep2_peakome.txt", sep = "\t", header = TRUE))
counts <- subset(df, select = c(6,7))
class(counts) <- "numeric"
counts <- as.matrix(counts)
counts <- as.data.frame(counts)
counts.log <- log2(counts)
plot(counts.log$cntrep1, counts.log$cntrep2)
ggscatter(counts.log, x = "cntrep1", y = "cntrep2", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "CUT&Tag replicate 1 before norm", ylab = "CUT&Tag replicate 2 before norm")
```
#We usually do this using TMM method implemented in edgeR package in R for normalization. I then plotted the log transformed normalized data in the form of a scatter plot

```{r pressure, echo=FALSE}
#Run edgeR to normalize the data using TMM
dge <- DGEList(counts=counts)
dge.norm <- calcNormFactors(dge, method="TMM") #TMM method for normalization
dge.norm$samples #to check the normalization factors and library size
y <- cpm(dge.norm,  normalized.lib.sizes = TRUE) #returns cpm while maintaining library size normalizations
```

#After having looked at the library size and normalization factors, lets log transform the data and look at the reproducibility
```{r pressure, echo=FALSE}
y <- as.data.frame(y) #Converting the list to a dataframe
y.log <- log2(y + 2) #transforming the data to the log scale for readability

plot(y.log$cntrep1, y.log$cntrep2, xlab = "CUT&Tag replicate 1", ylab = "CUT&Tag replicate 2") #Basic scatter plot

#Calculating Pearson Correlation Coefficient on the normalized data
pcc <- cor.test(y.log$cntrep1, y.log$cntrep2, method = "pearson")

#Including the PCC and p-value in the scatter plot. 
ggscatter(y.log, x = "cntrep1", y = "cntrep2", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "CUT&Tag replicate 1", ylab = "CUT&Tag replicate 2")
```
#I will next use the peakome generated for Merged CUT&Tag and Merged ChIP dataset against CUT&Tag_intersection_peaks (e-1 MACS setting). 
#Lets load the peakome dataset 
```{r}
#Loading the peakome data for the merged CUT&Tag and ChIP files
df1 <- as.matrix(read.table("CUTnTag_combined_k4me3chip_peakome.txt", sep = "\t", header = TRUE))
counts.merge <- subset(df1, select = c(6,7)) #Selecting the relevant columns for this analysis
class(counts.merge) <- "numeric"
counts.merge <- as.matrix(counts.merge)
counts.merge <- as.data.frame(counts.merge) #Converting the matrix to a data frame
```

#Lets look at the reproducibility between Cntmerge and ChIP merged data after log2 transforming the dataset
```{r}
counts.merge.log <- log2(counts.merge) #log transforming the data
plot(counts.merge.log$cntmerge, counts.merge.log$chipmerge) #Basic plot
#Adding more layers and calculating Pearson Correlation coefficient for the data
ggscatter(counts.merge.log, x = "cntmerge", y = "chipmerge", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "CUT&Tag merge before norm", ylab = "ChIP merge before norm")
```
#The PCC value is low suggesting low reproducibility between Cut&Tag merged and ChIP merged samples. Lets try the same thing after normalization
```{r}
dge.merge <- DGEList(counts=counts.merge)
dge.merge.norm <- calcNormFactors(dge.merge, method="TMM") #TMM method for normalization
dge.merge.norm$samples #to check the normalization factors and library size
y1 <- cpm(dge.merge.norm,  normalized.lib.sizes = TRUE) #returns cpm while maintaining library size normalizations
```


```{r}
y1 <- as.data.frame(y1) #Converting the list to a dataframe
y1.log <- log2(y1 + 2) #transforming the data to the log scale for readability
#Basic scatter plot
plot(y1.log$cntmerge, y1.log$chipmerge, xlab = "CUT&Tag merged", ylab = "CUT&Tag merged")
#Calculating Pearson Correlation Coefficient on the normalized data
pcc <- cor.test(y1.log$cntmerge, y1.log$chipmerge, method = "pearson")
ggscatter(y1.log, x = "cntmerge", y = "chipmerge", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "CUT&Tag merge", ylab = "H3K4me3 ChIP merge")
```
#There is low reproducibility between cnt merge and chip merged samples. It will be best to double-check whether I used the correct set of files for peakome generation. I checked and actually used the correct files for peakome. We will have to discuss more why this is the case!

