---
title: "Cut&Tag_DataExploration"
author: "Aditya Mahadevan"
date: "1/30/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---
Since we got diff types of distribution of peak overlaps with CUT&Tag and ChIP comparisons, I would like to plot the overlap in a different manner using GRanges and bed intersect tool in R.
#Load the packages
```{r setup, include=FALSE}
library(edgeR)
library(tidyverse)
library(GenomicRanges)
library(ggpubr)
```
#Setting up the options
```{r}
options(stringsAsFactors = FALSE)
def.par <- par(no.readonly = T)
```

#Function to intersect two bedfiles using GRanges
```{r}
intersect_bed <- function(x, y){
  # load into GRanges object
  a <- makeGRangesFromDataFrame(x, keep.extra.columns = T)
  b <- makeGRangesFromDataFrame(y, keep.extra.columns = T)
  # find overlaps
  my_hit <- findOverlaps(a, b)
  my_df  <- data.frame(b[subjectHits(my_hit)])
}
```
#Load the data of interest
```{r}
x <- read.delim("CUTnTag_merged_k4me3chip_mm10_peakome.txt")
x <- x %>%
  mutate(id = paste0("peak",seq.int(nrow(x)))) %>%
    select(chr, start, end, id, cntmerge, chipmerge)

```

#Perform DE analysis and normalize using edgeR TMM normalization method
```{r}
### Asign groups to replicates
group <- factor(c(1,2))
y <- DGEList(counts=x[,5:6], group=group)
### TMM normalization
y <- calcNormFactors(y)
y$samples
```

#Plot the data after log2 transformation and export the data for visualization
```{r}
data.cpm <- as.data.frame(cpm(y, normalized.lib.sizes = TRUE))
data <- round(log2(data.cpm + 1), 4)
data <- cbind(x[1:4],data) 
keep <- rowSums(cpm(y)>1) >= 2
data.filter <- data[keep, ]
### Visualize batch effects using MDS in edgeR
#plotMDS(data.filter)
```

#Use edgeR to calculate differential H3K4me3

#There is no dispersion as there are only one set of samples
```{r}
y <- estimateDisp(y) # after calcNormFactors; Dispersion is NA

#y <- estimateTagwiseDisp(y)
# exact test
#et <- exactTest(y)
# export table
#df <- as.data.frame(data.filter)
#data.filter$adjP <- p.adjust(df$PValue, method = "BH")
#df.join <- cbind(x, df)
## filter peaks with RPM > 1 -- filtering peaks after
#df.filter <- data.filter[keep, ]
```

#Filtering the data further (not required for this analysis)
```{r}
#nrow(df.filter %>% filter(logFC > 1 & adjP < .01)) #73
#nrow(df.filter %>% filter(logFC < -1 & adjP < .01)) #11812
```

#Defining the CAST hotspots
```{r}
# CAST hotspots defined by unique H3K4me3 sites
unique <- "E_binding_sites_R.mm10.bed"
# get hotspot annotations
unique <- read.delim((unique), header = F)
names(unique) <- c('chr', 'start', 'end', 'peak', 'score', 'strand')
```

#Defining the TSS (Gencode)
```{r}
# promoter annotation using ChromHMM state 1
tss <- "mm10_gencode_tss_unique.bed"
# get TSS annotations
tss <- read.delim((tss), header = F)
names(tss) <- c('chr', 'start', 'end')
```

## Intersect data with CAST hotspots
```{r}
df.b6 <- intersect_bed(unique, data.filter) %>% distinct(id, .keep_all = T)
df.notb6 <- subset(data.filter, !(id %in% df.b6$id))
#df.b6 <- subset(df.b6, !(start %in% df.fp$id))
```

#Intersect the data similarly for TSS
```{r}
df.tss <- intersect_bed(tss, data.filter) %>% distinct(id, .keep_all = T)
df.nottss <- subset(data.filter, !(id %in% df.tss$id))
#df.tss.f <- subset(df.tss, !(id %in% state1.hotspots$id))
```

#Making the MA plot for CAST hotspots
```{r}
#png("MA_plot_cntmerge_chipmerge_at_CASThotspots.png", width=6, height=6, units="in", res=600)
plot(df.notb6$cnt ~ df.notb6$chip,
     type = "n",
     #ylim=c(-6,6), xlim=c(-1,10), 
     ann = T)
grid (NULL,NULL, lty = 3, col = "lightgrey")
points(df.notb6$cnt ~ df.notb6$chip, 
       pch=20, 
       col=rgb(0,0,0,0.25))
points(df.b6$cnt ~ df.b6$chip, 
       pch=20, 
       col=rgb(40/250,170/250,250/250,0.25))
pcc.cast <- cor.test(df.notb6$cnt, df.notb6$chip, method = "pearson")
#abline(h = 1, col = "red")
#abline(h = -1, col = "red")
#dev.off()
```

#Making the MA plot for TSS
```{r}
#png("./MA_plot_cntmerge_chipmerge_at_TSS", width=6, height=6, units="in", res=600)
plot(df.nottss$cnt ~ df.nottss$chip,
     type = "n",
     #ylim=c(-6,6), xlim=c(-1,10), 
     ann = T)
grid (NULL,NULL, lty = 3, col = "lightgrey")
points(df.nottss$cnt ~ df.nottss$chip, 
       pch=20, 
       col=rgb(0,0,0,0.25))
points(df.tss$cnt ~ df.tss$chip, 
       pch=20, 
       col=rgb(40/250,170/250,250/250,0.25))
pcc_tss <- cor.test(df.nottss$cnt, df.nottss$chip, method = "pearson")
#dev.off()
```


```{r}
ggscatter(df.b6, x = "chipmerge", y = "cntmerge", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "CUT&Tag at hotspots", ylab = "ChIP at hotspots")
```


```{r}
ggscatter(df.tss, x = "chipmerge", y = "cntmerge", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "CUT&Tag at TSS", ylab = "ChIP at TSS")
```
